---
title: "Growth Rate Charts"
output:
  html_document:
    self_contained: false
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(plotly)
library(scales)
library(RColorBrewer)
library(tidyr)
library(knitr)

# Load saved dataframes
load("growth_dfs/GDP.RData")
load("growth_dfs/FDI.RData")
load("growth_dfs/GFCF.RData")
load("growth_dfs/trade.RData")
load("growth_dfs/debt.RData")
load("growth_dfs/debt2.RData")
load("growth_dfs/combined.RData")

line_growth_rate <- function(data, growth_rate="growth_rate", year = "year", country = "country", title = "Growth Rate Over Time") {
  # Convert growth_rate column to symbol for dynamic column selection
  growth_rate <- rlang::sym(growth_rate)
  year <- rlang::sym(year)
  country <- rlang::sym(country)
  
  ggplot(data, aes(x = !!year, y = !!growth_rate, colour = !!country)) +
    geom_line(linewidth = 1.2) +  # Make lines thicker for better readability
    scale_color_brewer(palette = "Set1") +  # Use a colorblind-friendly palette
    scale_y_continuous(labels = percent_format(scale = 1)) +  # Format y-axis as percentages
    labs(
      title = title,
      x = "Year",
      y = "Growth Rate (%)",
      colour = "Country"
    ) +
    theme_minimal(base_size = 14) +  # Set a minimal theme with a larger font size
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),  # Center and bold title
      axis.title.x = element_text(margin = margin(t = 10)),  # Add margin to x-axis title
      axis.title.y = element_text(margin = margin(r = 10)),  # Add margin to y-axis title
      axis.text = element_text(size = 12),  # Increase axis text size for readability
      legend.position = "bottom",  # Place the legend at the bottom for better use of space
      legend.title = element_text(face = "bold")  # Bold legend title
    )
}

bar_growth_rate <- function(data, growth_rate="growth_rate", country = "country", year = "year", title = "GDP Growth Rate") {
  world_colour <- brewer.pal(n = 3, "Set1")[3]
  ggplot(data %>%
           filter(!!rlang::sym(country) == "World"), 
         aes(x = factor(!!rlang::sym(year)), y = !!rlang::sym(growth_rate), fill = !!rlang::sym(country))) + 
    geom_col(position = "dodge", width = 0.7) +  # Use dodged bars for country comparison
    scale_fill_manual(values = c("World" = world_colour)) +
    scale_y_continuous(labels = percent_format(scale = 1)) +  # Format y-axis as percentages
    labs(
      title = title,
      x = "Year",
      y = "Growth Rate (%)",
      fill = "Country"
    ) +
    theme_minimal(base_size = 14) +  # Minimal theme with increased font size
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),  # Center and bold title
      axis.title.x = element_text(margin = margin(t = 10)),  # Add space to x-axis title
      axis.title.y = element_text(margin = margin(r = 10)),  # Add space to y-axis title
      axis.text.x = element_text(angle = 45, hjust = 1, size = 10),  # Tilt x-axis labels for readability
      axis.text.y = element_text(size = 12),  # Increase y-axis label size
      legend.position = "bottom",  # Place legend at the bottom
      legend.title = element_text(face = "bold")  # Bold legend title
    )
}

legend_bottom <- list(
  orientation = "h",  # Horizontal legend
  x = 0.5,            # Center horizontally
  xanchor = "center", # Align the center
  y = -0.2,            # Place it below the plot,
  title = list(text = NULL)
)

annualized_growth_rate <- function(data, start_year, end_year, growth_col = "GDP_2015") { #data = combined
  # Ensure the data is filtered to the relevant years
  data_filtered <- data %>% 
    filter(year %in% c(start_year, end_year))
  
  if (nrow(data_filtered) != 2) {
    stop("Dataframe has too many or too few rows for those years.")
  }
  
  # Extract GDP values for the start and end years
  gdp_start <- data_filtered %>% filter(year == start_year) %>% pull(!!rlang::sym(growth_col))
  gdp_end <- data_filtered %>% filter(year == end_year) %>% pull(!!rlang::sym(growth_col))
  
  # Calculate the number of years
  num_years <- end_year - start_year
  
  # Calculate the annualized growth rate
  growth_rate <- (gdp_end / gdp_start)^(1 / num_years) - 1
  
  return(paste0(round(growth_rate*100,1), "%"))
}

average_growth_rate <- function(data, start_year, end_year, growth_col = "GDP_2015_gr") {
  # Ensure the data is filtered to the relevant years
  data_filtered <- data %>% 
    filter(year >= start_year & year <= end_year)
  
  if (nrow(data_filtered) == 0) {
    stop("No data available for the specified range of years.")
  }
  
  # Extract the specified growth rate column and calculate the average
  avg_growth_rate <- data_filtered %>% 
    summarise(avg_growth = mean(!!rlang::sym(growth_col), na.rm = TRUE)) %>% 
    pull(avg_growth)
  
  # Return the result as a percentage string
  return(paste0(round(avg_growth_rate, 1), "%"))
}
```

Reminder: before publishing anything Ed or someone else should check we are dealing with missing data in the correct way. e.g. in some datasets combining multiple countries into developing or developed when for early years some countries' data are missing. 


# GDP, FDI, and trade {.tabset}

- For this plot GDP is real (2015 US$) pulled straight from UNCTADstat, 
- FDI and trade are converted to (real 2015 US$) using the implied cpi from UNCTAD_stat for the world extracted from UNCTAD GDP data. Maybe we can discuss whether or not this is the best strategy. But I thought it better than putting growth in nominal variables and growth in real variables in the same chart. 

# Set a year as base year

<iframe 
  src="https://bakered91.shinyapps.io/growth_charts1/" 
  width="100%" 
  height="600px" 
  style="border: none;">
</iframe>

# Growth Rate {.tabset}

## Developing Economies 
```{r, echo=FALSE}
data_long <- combined %>%
  filter(year > 1997) %>% 
  filter(country == "Developing economies") %>% 
  gather(key = "growth_rate_type", value = "growth_rate_value", all_of(c("GDP_2015_gr", "FDI_2015_gr", "trade_2015_gr")))

d_ing = ggplot(data_long, aes(x = year, y = growth_rate_value, colour = growth_rate_type)) +
  geom_line(linewidth = 1.2) +  # Make lines thicker for better readability
  scale_color_brewer(palette = "Set1") +  # Color palette
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
  labs(
    title = "Growth Rate of Developing Economies",
    x = "Year",
    y = "Growth Rate (%)",
    colour = NULL
  ) +
  theme_minimal(base_size = 14) +  # Minimal theme with larger font size
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),  # Center and bold title
    axis.title.x = element_text(margin = margin(t = 10)),  # Add margin to x-axis title
    axis.title.y = element_text(margin = margin(r = 10)),  # Add margin to y-axis title
    axis.text = element_text(size = 12),  # Increase axis text size for readability
    legend.position = "bottom",  # Place the legend at the bottom for better use of space
    legend.title = element_text(face = "bold")  # Bold legend title
  )

# Use plotly for interactive plotting
d_ing_plotly <- ggplotly(d_ing) %>%
  layout(legend = legend_bottom)
d_ing_plotly 

```


## World

```{r, echo=FALSE}
data_long <- combined %>%
  filter(year > 1997) %>% 
  filter(country == "World") %>% 
  gather(key = "growth_rate_type", value = "growth_rate_value", all_of(c("GDP_2015_gr", "FDI_2015_gr", "trade_2015_gr")))

d_ing = ggplot(data_long, aes(x = year, y = growth_rate_value, colour = growth_rate_type)) +
  geom_line(linewidth = 1.2) +  # Make lines thicker for better readability
  scale_color_brewer(palette = "Set1") +  # Color palette
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
  labs(
    title = "Growth Rate of World",
    x = "Year",
    y = "Growth Rate (%)",
    colour = NULL
  ) +
  theme_minimal(base_size = 14) +  # Minimal theme with larger font size
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),  # Center and bold title
    axis.title.x = element_text(margin = margin(t = 10)),  # Add margin to x-axis title
    axis.title.y = element_text(margin = margin(r = 10)),  # Add margin to y-axis title
    axis.text = element_text(size = 12),  # Increase axis text size for readability
    legend.position = "bottom",  # Place the legend at the bottom for better use of space
    legend.title = element_text(face = "bold")  # Bold legend title
  )

# Use plotly for interactive plotting
d_ing_plotly <- ggplotly(d_ing) %>%
  layout(legend = legend_bottom)
d_ing_plotly 

```



## Developed Economies 

```{r, echo=FALSE}
data_long <- combined %>%
  filter(year > 1997) %>% 
  filter(country == "Developed economies") %>% 
  gather(key = "growth_rate_type", value = "growth_rate_value", all_of(c("GDP_2015_gr", "FDI_2015_gr", "trade_2015_gr")))

d_ing = ggplot(data_long, aes(x = year, y = growth_rate_value, colour = growth_rate_type)) +
  geom_line(linewidth = 1.2) +  # Make lines thicker for better readability
  scale_color_brewer(palette = "Set1") +  # Color palette
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentages
  labs(
    title = "Growth Rate of Developed economies",
    x = "Year",
    y = "Growth Rate (%)",
    colour = NULL
  ) +
  theme_minimal(base_size = 14) +  # Minimal theme with larger font size
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),  # Center and bold title
    axis.title.x = element_text(margin = margin(t = 10)),  # Add margin to x-axis title
    axis.title.y = element_text(margin = margin(r = 10)),  # Add margin to y-axis title
    axis.text = element_text(size = 12),  # Increase axis text size for readability
    legend.position = "bottom",  # Place the legend at the bottom for better use of space
    legend.title = element_text(face = "bold")  # Bold legend title
  )

# Use plotly for interactive plotting
d_ing_plotly <- ggplotly(d_ing) %>%
  layout(legend = legend_bottom)
d_ing_plotly 

```



# Growth Rates Tables {.tabset}

## Average Annual Growth Rates {.tabset}

### World 
```{r, echo=FALSE}

year_breaks = c(1998, 2008, 2018, 2023)

generate_growth_table <- function(row_names, year_breaks, region, data) {
  # Create column names dynamically based on year_breaks
  colnames_growth_table <- paste0(year_breaks[-length(year_breaks)] + 1, 
                                  "-", 
                                  year_breaks[-1])
  
  # Generate the growth table dynamically
  growth_table <- data.frame(
    row.names = row_names,
    setNames(
      lapply(1:(length(year_breaks) - 1), function(i) {
        sapply(row_names, function(col) 
          average_growth_rate(data %>% filter(country == region), 
                              year_breaks[i] + 1, year_breaks[i + 1], 
                              sym(col)))
      }),
      colnames_growth_table
    ),
    check.names = FALSE
  )
  
  return(growth_table)
}


growth_table = generate_growth_table(row_names = c("GDP_2015_gr", "FDI_2015_gr", "trade_2015_gr"), 
                      year_breaks = year_breaks, 
                      region = "World", 
                      data = combined)

# Render the table in RMarkdown
kable(growth_table)

```

### Developing economies
```{r, echo=FALSE}
growth_table = generate_growth_table(row_names = c("GDP_2015_gr", "FDI_2015_gr", "trade_2015_gr"), 
                      year_breaks = year_breaks, 
                      region = "Developing economies", 
                      data = combined)

# Render the table in RMarkdown
kable(growth_table)

```

### Developed economies
```{r, echo=FALSE}
growth_table = generate_growth_table(row_names = c("GDP_2015_gr", "FDI_2015_gr", "trade_2015_gr"), 
                      year_breaks = year_breaks, 
                      region = "Developed economies", 
                      data = combined)

# Render the table in RMarkdown
kable(growth_table)

```


## Anualised Growth Rates {.tabset}

### World 
```{r, echo=FALSE}
generate_annualized_growth_table <- function(row_names, year_breaks, region, data) {
  # Create column names dynamically based on year_breaks
  colnames_growth_table <- paste0(year_breaks[-length(year_breaks)], 
                                  "-", 
                                  year_breaks[-1])
  
  # Generate the growth table dynamically
  growth_table <- data.frame(
    row.names = row_names,
    setNames(
      lapply(1:(length(year_breaks) - 1), function(i) {
        sapply(row_names, function(col) 
          annualized_growth_rate(data %>% filter(country == region), 
                                 year_breaks[i], year_breaks[i + 1], 
                                 sym(col)))
      }),
      colnames_growth_table
    ),
    check.names = FALSE
  )
  
  return(growth_table)
}

growth_table = generate_annualized_growth_table(row_names = c("GDP_2015", "FDI_2015", "trade_2015"), 
                      year_breaks = year_breaks, 
                      region = "World", 
                      data = combined)

# Render the table in RMarkdown
kable(growth_table)

```

### Developing economies
```{r, echo=FALSE}

growth_table = generate_annualized_growth_table(row_names = c("GDP_2015", "FDI_2015", "trade_2015"), 
                      year_breaks = year_breaks, 
                      region = "Developing economies", 
                      data = combined)

# Render the table in RMarkdown
kable(growth_table)

```

### Developed economies
```{r, echo=FALSE}
growth_table = generate_annualized_growth_table(row_names = c("GDP_2015", "FDI_2015", "trade_2015"), 
                      year_breaks = year_breaks, 
                      region = "Developed economies", 
                      data = combined)

# Render the table in RMarkdown
kable(growth_table)

```


# Growth Rates Tables (per capita) {.tabset}

## Average Annual Growth Rates {.tabset}

### World 
```{r, echo=FALSE}
growth_table = generate_growth_table(row_names = c("GDP_2015_pc_gr", "FDI_2015_pc_gr", "trade_2015_pc_gr"), 
                      year_breaks = year_breaks, 
                      region = "World", 
                      data = combined)

# Render the table in RMarkdown
kable(growth_table)

```

### Developing economies
```{r, echo=FALSE}
growth_table = generate_growth_table(row_names = c("GDP_2015_pc_gr", "FDI_2015_pc_gr", "trade_2015_pc_gr"), 
                      year_breaks = year_breaks, 
                      region = "Developing economies", 
                      data = combined)

# Render the table in RMarkdown
kable(growth_table)

```

### Developed economies
```{r, echo=FALSE}
growth_table = generate_growth_table(row_names = c("GDP_2015_pc_gr", "FDI_2015_pc_gr", "trade_2015_pc_gr"), 
                      year_breaks = year_breaks, 
                      region = "Developed economies", 
                      data = combined)

# Render the table in RMarkdown
kable(growth_table)

```


## Anualised Growth Rates {.tabset}

### World 
```{r, echo=FALSE}
growth_table = generate_annualized_growth_table(row_names = c("GDP_2015_pc", "FDI_2015_pc", "trade_2015_pc"), 
                      year_breaks = year_breaks, 
                      region = "World", 
                      data = combined)

# Render the table in RMarkdown
kable(growth_table)

```

### Developing economies
```{r, echo=FALSE}

growth_table = generate_annualized_growth_table(row_names = c("GDP_2015_pc", "FDI_2015_pc", "trade_2015_pc"), 
                      year_breaks = year_breaks, 
                      region = "Developing economies", 
                      data = combined)

# Render the table in RMarkdown
kable(growth_table)

```

### Developed economies
```{r, echo=FALSE}
growth_table = generate_annualized_growth_table(row_names = c("GDP_2015_pc", "FDI_2015_pc", "trade_2015_pc"), 
                      year_breaks = year_breaks, 
                      region = "Developed economies", 
                      data = combined)

# Render the table in RMarkdown
kable(growth_table)

```


# GDP Growth Rate {.tabset}

- UNCTADstat
- growth rate of real GDP (2015 US dollars) at market exchange rates

## Line
```{r, echo=FALSE}
# Filter data and plot the line chart for GDP Growth Rate
gdp_line <- GDP %>% 
  filter(year > 1997 & year < 2024) %>% 
  line_growth_rate(growth_rate = "GDP_2015_gr", title = "GDP Growth Rate")

# Use plotly for interactive plotting
gdp_line_plotly <- ggplotly(gdp_line) %>%
  layout(legend = legend_bottom)
gdp_line_plotly 
```

## Bar
```{r, echo=FALSE}
# Filter data and plot the bar chart for GDP Growth Rate
gdp_bar <- GDP %>% 
  filter(year > 1997 & year < 2024)  %>%
  bar_growth_rate(growth_rate = "GDP_2015_gr", title = "GDP Growth Rate")

# Use plotly for interactive plotting
gdp_bar_plotly <- ggplotly(gdp_bar)  %>%
  layout(legend = legend_bottom)
gdp_bar_plotly 
```

# FDI Growth Rate {.tabset}

- UNCTADstat
- FDI flow inwards, US current dollars

## Line
```{r, echo=FALSE}
# Filter data and plot the line chart for FDI Growth Rate
fdi_line <- FDI %>% 
  filter(year > 1997 & year < 2024)  %>% 
  line_growth_rate(growth_rate = "FDI_gr", title = "FDI Growth Rate")

# Use plotly for interactive plotting
fdi_line_plotly <- ggplotly(fdi_line) %>%
  layout(legend = legend_bottom)
fdi_line_plotly
```

## Bar
```{r, echo=FALSE}
# Filter data and plot the bar chart for FDI Growth Rate
fdi_bar <- FDI %>% 
  filter(year > 1997 & year < 2024) %>% 
  bar_growth_rate(growth_rate = "FDI_gr", title = "FDI Growth Rate")

# Use plotly for interactive plotting
fdi_bar_plotly <- ggplotly(fdi_bar)%>%
  layout(legend = legend_bottom)
fdi_bar_plotly
```

# GFCF Growth Rate {.tabset}

- world bank WDI (NE.GDI.FTOT.CD) 
- growth rate of current dollars 
- using world banks own world classification, but summing up over countries for developing and developed

## Line
```{r, echo=FALSE}
# Filter data and plot the line chart for GFCF Growth Rate
gfcf_line <- GFCF %>% 
  filter(year > 1997 & year < 2024) %>% 
  line_growth_rate(growth_rate = "GFCF_2015_gr", title = "GFCF Growth Rate")

# Use plotly for interactive plotting
gfcf_line_plotly <- ggplotly(gfcf_line) %>%
  layout(legend = legend_bottom)
gfcf_line_plotly
```

## Bar
```{r, echo=FALSE}
# Filter data and plot the bar chart for GFCF Growth Rate
gfcf_bar <- GFCF %>% 
  filter(year > 1997 & year < 2024)  %>%
  bar_growth_rate(growth_rate = "GFCF_2015_gr", title = "GFCF Growth Rate")

# Use plotly for interactive plotting
gfcf_bar_plotly <- ggplotly(gfcf_bar)%>%
  layout(legend = legend_bottom)
gfcf_bar_plotly
```

# Trade Growth Rate {.tabset}

- UNCTADstat
- growth rate of exports + imports

## Line current US $
```{r, echo=FALSE}
# Filter data and plot the line chart for Trade Growth Rate
trade_line <- trade %>% 
  filter(year > 1997 & year < 2024) %>% 
  line_growth_rate(growth_rate = "trade_gr", title = "Trade Growth Rate")

# Use plotly for interactive plotting
trade_line_plotly <- ggplotly(trade_line) %>%
  layout(legend = legend_bottom)
trade_line_plotly
```

## Bar current US $
```{r, echo=FALSE}
# Filter data and plot the bar chart for Trade Growth Rate
trade_bar <- trade %>% 
  filter(year > 1997 & year < 2024)  %>%
  bar_growth_rate(growth_rate = "trade_gr", title = "Trade Growth Rate")

# Use plotly for interactive plotting
trade_bar_plotly <- ggplotly(trade_bar)%>%
  layout(legend = legend_bottom)
trade_bar_plotly
```

## Line 2015 US $
```{r, echo=FALSE}
# Filter data and plot the line chart for Trade Growth Rate
trade_line <- trade %>% 
  filter(year > 1997 & year < 2024) %>% 
  line_growth_rate(growth_rate = "trade_2015_gr", title = "Trade Growth Rate")

# Use plotly for interactive plotting
trade_line_plotly <- ggplotly(trade_line) %>%
  layout(legend = legend_bottom)
trade_line_plotly
```

## Bar 2015 US $
```{r, echo=FALSE}
# Filter data and plot the bar chart for Trade Growth Rate
trade_bar <- trade %>% 
  filter(year > 1997 & year < 2024)  %>%
  bar_growth_rate(growth_rate = "trade_2015_gr", title = "Trade Growth Rate")

# Use plotly for interactive plotting
trade_bar_plotly <- ggplotly(trade_bar)%>%
  layout(legend = legend_bottom)
trade_bar_plotly

```


# Debt (as % of GDP) Growth Rate {.tabset}

- from world bank WDI
- growth rate of debt as a % of GDP
- World is world banks own category, developing and developed are the sum of all developing and developed countries, thats why it goes back further, and need to check how many countries it even hasvalues for. 

## Line
```{r, echo=FALSE}
# Filter data and plot the line chart for Debt Growth Rate
debt_line <- debt %>% 
  filter(year > 1997 & year < 2024)  %>% 
  line_growth_rate(growth_rate = "debt_gr", title = "Debt Growth Rate")

# Use plotly for interactive plotting
debt_line_plotly <- ggplotly(debt_line) %>%
  layout(legend = legend_bottom)
debt_line_plotly
```

## Bar
```{r, echo=FALSE}
# Filter data and plot the bar chart for Debt Growth Rate
debt_bar <- debt %>% 
  filter(year > 1997 & year < 2024) %>%
  bar_growth_rate(growth_rate = "debt_gr", title = "Debt Growth Rate")

# Use plotly for interactive plotting
debt_bar_plotly <- ggplotly(debt_bar)%>%
  layout(legend = legend_bottom)
debt_bar_plotly
```

# Debt Growth Rate (World of Debt) {.tabset}

- from world of debt, ultimately from IMF data. 
- growth rate of debt in US$
- values got by adding individual countries together to make the group. need to check if countries are missing. 

## Line
```{r, echo=FALSE}
# Filter data and plot the line chart for Debt Growth Rate
debt_line <- debt2 %>% 
  filter(year > 1997 & year < 2022)  %>% 
  line_growth_rate(growth_rate = "debt2_gr", title = "Debt Growth Rate") 
debt_line = debt_line +
  scale_x_continuous(limits = c(2000, NA))

# Use plotly for interactive plotting
debt_line_plotly <- ggplotly(debt_line) %>%
  layout(legend = legend_bottom)
debt_line_plotly
```

## Bar
```{r, echo=FALSE}
# Filter data and plot the bar chart for Debt Growth Rate
debt_bar <- debt2 %>% 
  filter(year > 1997 & year < 2024) %>%
  bar_growth_rate(growth_rate = "debt2_gr", title = "Debt Growth Rate") 
debt_bar = debt_bar +
  scale_x_discrete(limits = as.character(2000:2022))


# Use plotly for interactive plotting
debt_bar_plotly <- ggplotly(debt_bar)%>%
  layout(legend = legend_bottom)
debt_bar_plotly
```
